\textsf{Are large language models superhuman chemists?}

### 2024-03-27 src/tex/authors.tex
Update authors.tex
update authors with latest feedback
update authors
sync with overleaf
[showyourwork]
update authors and acknowledgements
update correlation plot and main text
[showyourwork] overleaf sync
add JCSM

% Question contributors
\credit{Benedict~Emoekabu}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
\credit{Aswanth~Krishnan}{0,0,0,0,0,0,0,0,1,0,0,0,0,0}
\credit{Tanya~Gupta}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
\credit{Mara~Wilhemli}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}
\credit{Macjonathan~Okereke}{0,1,0,0,1,0,0,0,0,0,0,0,0,0}

% Question answerers
\credit{Mehrdad Asgari}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
\credit{Juliane Eberhardt}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
\credit{Amir Elahi}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
\credit{Christina Glaubitz}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
\credit{Maximilian Greiner}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
\credit{Caroline Holick}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
\credit{Tim Hoffmann}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
\credit{Lea~C.~Klepsch}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
\credit{Yannik Köster}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
\credit{Fabian Alexander Kreth}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
\credit{Jakob~Meyer}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
\credit{Jan Peschel}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
\credit{Michael Ringleb}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
\credit{Nicole Roesner}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
\credit{Johanna Schreiber}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}
\credit{Ulrich~S.~Schubert}{0,0,0,1,0,0,0,0,0,0,0,0,0,0}
\credit{Leanne M. Stafast}{0,0,0,1,0,0,0,0,0,0,0,0,0,0}
\credit{Dinga Wonanke}{0,1,0,0,0,0,0,0,0,0,0,0,0,0}

% senior authors 
\credit{Michael~Pieler}{1,0,0,1,0,1,0,1,0,0,0,0,0,1}
\credit{Philippe~Schwaller}{1,1,0,1,1,1,0,0,0,1,0,0,0,0}
\credit{Kevin Maik Jablonka}{1,1,1,1,1,1,1,1,1,1,1,1,1,1}

\author[1,$\star$]{Adrian~Mirza~\orcidlink{0000-0003-4033-4235}}
\author[1,$\star$]{Nawaf~Alampara~\orcidlink{0009-0001-7697-7315}}
\author[1,$\star$]{Sreekanth~Kunchapu~\orcidlink{0009-0003-5752-0154}}
\author[]{Benedict~Emoekabu}
\author[4]{Aswanth~Krishnan~\orcidlink{0009-0008-2703-5613}}
\author[5,6]{Tanya~Gupta}
\author[1]{Mara~Willhelmi}
\author[1]{Macjonathan~Okereke~\orcidlink{0009-0007-1013-0502}}

\author[7]{Mehrdad~Asgari~\orcidlink{0000-0002-5427-1610}}
\author[13]{Juliane~Eberhardt~\orcidlink{0009-0000-3991-0704}}
\author[8]{Amir~Elahi}
\author[]{Christina~Glaubitz~\orcidlink{0000-0003-3412-0548}}
\author[1]{Maximilian~Greiner}
\author[1]{Caroline~Holick}
\author[1]{Tim~Hoffmann}
\author[1]{Lea~C.~Klepsch~\orcidlink{0009-0009-3849-1670}}
\author[1]{Yannik~Köster~\orcidlink{0000-0002-9125-3067}}
\author[2, 9]{Fabian~Alexander~Kreth~\orcidlink{0000-0002-5968-8706}}
\author[1]{Jakob~Meyer}
\author[1]{Jan~Peschel~\orcidlink{0009-0002-4787-2757}}
\author[1]{Michael~Ringleb~\orcidlink{0000-0002-7320-8529}}
\author[1]{Nicole~Roesner}

\author[1]{Johanna Schreiber}
\author[1]{Ulrich~S.~Schubert~\orcidlink{0000-0003-4978-4670}}
\author[1]{Leanne~M.~Stafast}
\author[10]{Dinga~Wonanke~\orcidlink{0000-0002-9066-2715}}

% Senior authors
\author[11,12]{Michael Pieler~\orcidlink{0000-0001-9186-7045}}
\author[5,6]{Philippe~Schwaller~\orcidlink{0000-0003-3046-6576}}
\author[1,2, 3, \Letter]{Kevin~Maik~Jablonka~\orcidlink{0000-0003-4894-4660}}

\affil[1]{Laboratory of Organic and Macromolecular Chemistry (IOMC), Friedrich Schiller University Jena, Humboldtstrasse 10, 07743 Jena, Germany}
\affil[2]{Center for Energy and Environmental Chemistry Jena (CEEC Jena), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
\affil[3]{Helmholtz Institute for Polymers in Energy Applications Jena (HIPOLE Jena), Lessingstrasse 12-14, 07743 Jena, Germany}

\affil[5]{Laboratory of Artificial Chemical Intelligence (LIAC), Institut des Sciences et Ing\'{e}nierie Chimiques, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Lausanne, Switzerland}
\affil[6]{National Centre of Competence in Research (NCCR) Catalysis, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Lausanne, Switzerland}

\affil[8]{Laboratory of Molecular Simulation (LSMO), Institut des Sciences et Ing\'{e}nierie Chimiques, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Sion, Switzerland}
\affil[10]{Theoretical Chemistry, Technische Universität Dresden, Dresden 01062, Germany}
\affil[11]{OpenBioML.org}
\affil[12]{Stability.AI}
\affil[9]{Institute for Technical Chemistry and Environmental Chemistry (ITUC), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
\affil[4]{QpiVolta Technologies Pvt Ltd}
\affil[7]{Department of Chemical Engineering \& Biotechnology, University of Cambridge, Philippa Fawcett Drive, Cambridge CB3 0AS, United Kingdom}
\affil[13]{Macromolecular Chemistry, University of Bayreuth, 95447 Bayreuth, Germany}
\affil[\Letter]{\texttt{mail@kjablonka.com}}
\affil[$\star$]{These authors contributed equally.}

>>>>

% Question contributors
\credit{Benedict~Emoekabu}{0,1,0,0,1,0,0,0,0,0,0,0,0,1}
\credit{Aswanth~Krishnan}{0,0,0,0,0,0,0,0,1,0,0,0,0,1}
\credit{Tanya~Gupta}{0,1,0,0,1,0,0,0,0,0,0,0,0,1}
\credit{Mara~Wilhelmi}{0,1,0,0,1,0,0,0,0,0,0,0,0,1}
\credit{Macjonathan~Okereke}{0,1,0,0,1,0,0,0,0,0,0,0,0,1}

% Question answerers
\credit{Mehrdad Asgari}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
\credit{Juliane Eberhardt}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
\credit{Amir Elahi}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
\credit{Christina Glaubitz}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
\credit{Maximilian Greiner}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
\credit{Caroline T. Holick}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
\credit{Tim Hoffmann}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
\credit{Lea~C.~Klepsch}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
\credit{Yannik Köster}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
\credit{Fabian Alexander Kreth}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
\credit{Jakob~Meyer}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
\credit{Jan Matthias Peschel}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
\credit{Michael Ringleb}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
\credit{Nicole Roesner}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
\credit{Johanna Schreiber}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
\credit{Ulrich~S.~Schubert}{0,0,0,1,0,0,0,0,0,0,0,0,0,1}
\credit{Leanne M. Stafast}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}
\credit{Dinga Wonanke}{0,1,0,0,0,0,0,0,0,0,0,0,0,1}

% senior authors 
\credit{Michael~Pieler}{1,0,0,1,0,1,0,1,0,0,0,0,0,1}
\credit{Philippe~Schwaller}{1,1,0,1,1,1,0,0,0,1,0,0,0,1}
\credit{Kevin Maik Jablonka}{1,1,1,1,1,1,1,1,1,1,1,1,1,1}

\author[1,$\star$]{Adrian~Mirza~\orcidlink{0000-0003-4033-4235}}
\author[1,$\star$]{Nawaf~Alampara~\orcidlink{0009-0001-7697-7315}}
\author[1,$\star$]{Sreekanth~Kunchapu~\orcidlink{0009-0003-5752-0154}}
\author[ ]{Benedict~Emoekabu}
\author[2]{Aswanth~Krishnan~\orcidlink{0009-0008-2703-5613}}
\author[3,4]{Tanya~Gupta~\orcidlink{0009-0001-9523-3290}}
\author[1]{Mara~Wilhelmi}
\author[1]{Macjonathan~Okereke~\orcidlink{0009-0007-1013-0502}}

\author[5]{Mehrdad~Asgari~\orcidlink{0000-0002-5427-1610}}
\author[6]{Juliane~Eberhardt~\orcidlink{0009-0000-3991-0704}}
\author[7]{Amir~Elahi}
\author[]{Christina~Glaubitz~\orcidlink{0000-0003-3412-0548}}
\author[1]{Maximilian~Greiner}
\author[1]{Caroline~T.~Holick\orcidlink{0009-0000-1724-2725}}
\author[1]{Tim~Hoffmann~\orcidlink{0009-0004-0230-6115}}
\author[1]{Lea~C.~Klepsch~\orcidlink{0009-0009-3849-1670}}
\author[1]{Yannik~Köster~\orcidlink{0000-0002-9125-3067}}
\author[8, 9]{Fabian~Alexander~Kreth~\orcidlink{0000-0002-5968-8706}}
\author[1]{Jakob~Meyer}
\author[1]{Jan~Matthias~Peschel~\orcidlink{0009-0002-4787-2757}}
\author[1]{Michael~Ringleb~\orcidlink{0000-0002-7320-8529}}
\author[1, 10]{Nicole~Roesner~\orcidlink{0000-0002-5133-775X}}

\author[1, 10]{Johanna Schreiber~\orcidlink{0009-0000-0991-8967}}
\author[1,8,10, 11]{Ulrich~S.~Schubert~\orcidlink{0000-0003-4978-4670}}
\author[1]{Leanne~M.~Stafast}
\author[10]{Dinga~Wonanke~\orcidlink{0000-0002-9066-2715}}

% Senior authors
\author[12,13]{Michael Pieler~\orcidlink{0000-0001-9186-7045}}
\author[3,4]{Philippe~Schwaller~\orcidlink{0000-0003-3046-6576}}
\author[1,8,10, 11, \Letter]{Kevin~Maik~Jablonka~\orcidlink{0000-0003-4894-4660}}

\affil[1]{Laboratory of Organic and Macromolecular Chemistry (IOMC), Friedrich Schiller University Jena, Humboldtstrasse 10, 07743 Jena, Germany}
\affil[8]{Center for Energy and Environmental Chemistry Jena (CEEC Jena), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
\affil[11]{Helmholtz Institute for Polymers in Energy Applications Jena (HIPOLE Jena), Lessingstrasse 12-14, 07743 Jena, Germany}
\affil[10]{Jena Center for Soft Matter (JCSM), Friedrich Schiller University Jena, Philosophenweg 7, 07743 Jena, Germany}

\affil[3]{Laboratory of Artificial Chemical Intelligence (LIAC), Institut des Sciences et Ing\'{e}nierie Chimiques, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Lausanne, Switzerland}
\affil[4]{National Centre of Competence in Research (NCCR) Catalysis, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Lausanne, Switzerland}

\affil[7]{Laboratory of Molecular Simulation (LSMO), Institut des Sciences et Ing\'{e}nierie Chimiques, Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL), Sion, Switzerland}
\affil[10]{Theoretical Chemistry, Technische Universität Dresden, Dresden 01062, Germany}
\affil[12]{OpenBioML.org}
\affil[13]{Stability.AI}
\affil[9]{Institute for Technical Chemistry and Environmental Chemistry (ITUC), Friedrich Schiller University Jena, Philosophenweg 7a, 07743 Jena, Germany}
\affil[2]{QpiVolta Technologies Pvt Ltd}
\affil[5]{Department of Chemical Engineering \& Biotechnology, University of Cambridge, Philippa Fawcett Drive, Cambridge CB3 0AS, United Kingdom}
\affil[6]{Macromolecular Chemistry, University of Bayreuth, 95447 Bayreuth, Germany}
\affil[\Letter]{\texttt{mail@kjablonka.com}}
\affil[$\star$]{These authors contributed equally.}

### 2024-03-27 src/tex/ms.tex
add JCSM

In this work, we report a novel benchmarking framework, which we call \chembench, and use it to reveal limitations of current frontier models for use in the chemical sciences.
Our benchmark consists of \variable{output/total_number_of_questions.txt}\xspace question answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
Our corpus covers a large fraction of the topics taught in undergraduate and graduate chemistry curricula and can be used to evaluate any system that can return text (i.e., also tool-augmented systems).

@@

% Radar chart
\paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and classified all questions in the \chembench corpus into these topics based on hand-crafted rules as well as classifier models operating on the question texts.
We then computed the percentage of questions that the models, or humans, answered completely correctly for each topic.
\begin{figure}[!h]
    \centering
    \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus. The performance is measured as the fraction of questions that were answered completely correct by the models.
    The best score for every dimension is one (all questions answered correctly) and the worst score is zero (no question answered completely correctly). A larger colored area indicates a better performance.
    This figure shows the performance on the subset of questions that was answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
    }
    \label{fig:all_questions_models_completely_correct_radar_human}
    \script{analyze_model_reports.py}
\end{figure}
In this spider chart, the worst score for every dimension is zero (no question answered completely correctly) and the best score is one (all questions answered correctly). 
Thus, a larger colored area indicates a better performance. 
One can observe that this performance varies widely across models and topics. 
While macromolecular chemistry and biochemistry receive relatively high scores for many models, this is not the case for topics such as chemical safety or analytical chemistry.
In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4) while this question appeared easier (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans.
Importantly, the human experts are given a drawing of the compounds, whereas models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule). 
These findings also shine an interesting light on the value of textbook-inspired questions. 
A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students. 
On those questions the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
For instance, while the overall performance in the chemical safety topic is low, the models would pass the examination for a chemical license based on a subset of a questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for GPT-4, \variable{output/subset_scores/is_gfk_claude3.txt}\% for Claude 3, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.

% Calculation vs. no calculation
Given that some of the questions in \chembench require elaborate calculations, we also analyzed if the models perform differently on those questions.
For this purpose, we manually annotated the questions that require calculations and performed the analysis on only those questions.  % TODO add some analysis here

>>>>

In this work, we report a novel benchmarking framework  (\Cref{fig:overview_figure}), which we call \chembench, and use it to reveal limitations of current frontier models for use in the chemical sciences.
Our benchmark consists of \variable{output/total_number_of_questions.txt}\xspace question answer pairs manually (\variable{output/manually_generated.txt}\unskip) or semi-automatically (\variable{output/automatically_generated.txt}\unskip) compiled from diverse sources.
Our corpus covers a large fraction of the topics taught in undergraduate and graduate chemistry curricula and can be used to evaluate any system that can return text (i.e., also tool-augmented systems).

@@

% Radar chart
\paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and classified all questions in the \chembench corpus into these topics based on hand-crafted rules as well as classifier models operating on the question texts.
We then computed the percentage of questions that the models, or humans, answered completely correctly for each topic.
\begin{figure}[!h]
    \centering
    \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus. The performance is measured as the fraction of questions that were answered completely correct by the models.
    The best score for every dimension is one (all questions answered correctly) and the worst score is zero (no question answered completely correctly). A larger colored area indicates a better performance.
    This figure shows the performance on the subset of questions that was answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
    }
    \label{fig:all_questions_models_completely_correct_radar_human}
    \script{analyze_model_reports.py}
\end{figure}
In this spider chart, the worst score for every dimension is zero (no question answered completely correctly) and the best score is one (all questions answered correctly). 
Thus, a larger colored area indicates a better performance. 
One can observe that this performance varies widely across models and topics. 
While macromolecular chemistry and biochemistry receive relatively high scores for many models, this is not the case for topics such as chemical safety or analytical chemistry.
In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4) while this question appeared easier (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans.
Importantly, the human experts are given a drawing of the compounds, whereas models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule). 
These findings also shine an interesting light on the value of textbook-inspired questions. 
A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students. 
On those questions the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of a questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for GPT-4, \variable{output/subset_scores/is_gfk_claude3.txt}\% for Claude 3, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.

It is important to note that the model performance for some topics, however, is underestimated in the current evaluation. This is because models that are provided via \glspl{api} typically have safety mechanisms that prevent them from providing answers that the provider deems unsafe.
For instance, models might refuse to provide answers about cyanides. To overcome this, direct access to the model weights would be required, and we strive to collaborate with the developers of frontier models to overcome this limitation in the future.
